---
title: "p8105_hw2_yc4584"
author: "Yingyu Cui"
date: "2024-09-30"
output: github_document
---

```{r set up}
library(tidyverse)
```


# Problem 1: NYC Transit Data

Read and clean the data
```{r read and clean the data}
subway_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = "") |> 
  janitor::clean_names() |> 
  select(line:entry, vending, ada) |>
  mutate(entry = case_match(entry, 
                            "YES" ~ TRUE,
                            "NO" ~ FALSE))
```
For this data set, it contains the variables "line", "station_name",  "routes", "station_latitude", "station_longitude", "entrance_type", "entry", "vending" and "ada" columns. The "line", "station_name", "route1-7", "entrance_type" and "vending" are character variables, the "station_latitude", "station_longitude", "route8-11" are numeric variables, and the "entry" and "ada" are logical variables.
For data cleaning, I used the janitor package to clean the column names and the select function to select the columns I need. I also used the case_match function to convert the "entry" column to logical variables.
The dimension of the cleaned data set is 1868 rows and 19 columns.
Except the columns of routes belonging to different kinds of variables, I think the data set is pretty much clean.

Question1: 
```{r distinct stations}
distinct_stations = subway_df |> 
  distinct(line, station_name)
num_distinct_stations = nrow(distinct_stations)
print(num_distinct_stations)
```
There are 465 distinct stations in the data set.

Question2:
```{r ADA compliance}
num_ada_compliance = subway_df |> 
  filter(ada == "TRUE") |> 
  nrow()
print(num_ada_compliance)
```
Therefore, 468 stations are ADA compliant.

Question3:
```{r vending}
num_without_vending = subway_df |> 
  filter(vending == "NO" & entry == "TRUE") |> 
  nrow()
print(num_without_vending)
```
Therefore, the proportion is 69/1867 = 0.037.

Reformat data so that route number and route name are distinct variables.
```{r reformat data}
subway_df_reformat = subway_df |> 
  mutate(across(starts_with("route"), as.character)) |> 
  pivot_longer(cols = starts_with("route"),
               names_to = "route_number", 
               values_to = "route_name",
               values_drop_na = TRUE)
```
How many serve the A train?
```{r A train}
num_A_train = subway_df_reformat |> 
  filter(route_name == "A") |> 
  distinct(line, station_name) |> 
  nrow()
print(num_A_train)
```
Therefore,60 distinct stations serve the A train.

Of the stations that serve the A train, how many are ADA compliant?
```{r A train ADA}
num_A_train_ADA = subway_df_reformat |> 
  filter(route_name == "A" & ada == "TRUE") |> 
  distinct(line, station_name) |> 
  nrow()
print(num_A_train_ADA)
```
Therefore, 17 stations that serve the A train are ADA compliant.




